---
output: html_document
---
#Practiacal Machine Learning Prediction Assignment

##Synopsis
This assignment is to build a prediction model to apply in the prediction quiz. Two models will be built, a Decision Tree model and a Random Forest model, and evaluated. The model with the better accuracy in predicting on a validation set will be selected.

##Load R Packeges
```{r}
library(caret)
library(randomForest)
library(rpart)
library(rattle)
```

##Load the Data
```{r}
Training.Data <- read.csv("pml-training.csv")
Testing.Data <- read.csv("pml-testing.csv")
```
##Clean the Data
```{r}
#Remove the subject and timestap data
Training.Data <- Training.Data[,-(1:5)]
Testing.Data <- Testing.Data[,-(1:5)]

#Remove variables with near zero variance
Zero.Var <- nearZeroVar(Training.Data)
Training.Data <- Training.Data[,-Zero.Var]
Testing.Data <- Testing.Data[,-Zero.Var]

#Remove Variables that are all NAs
NAs <- colSums(is.na(Testing.Data))>5
Training.Data <- Training.Data[, NAs == FALSE]
Testing.Data <- Testing.Data[, NAs == FALSE]
```

##Partition the Training Data into a Training and a Validation Set
```{r}
set.seed(1316)
inTrain <- createDataPartition(Training.Data$classe, p = 0.6, list = FALSE)
Training <- Training.Data[inTrain,]
Validation <- Training.Data[-inTrain,]
```

##Random Forests Model
```{r}
set.seed(1316)
model.RF <- train(classe ~ ., data = Training, method = "rf", trControl=trainControl(method = "cv", number = 9), na.action = na.omit)
model.RF
varImp(model.RF)
```

##rpart Model
```{r}
set.seed(1316)
model.rpart <- train(classe ~ . , data = Training, method = "rpart", na.action = na.omit)
model.rpart
fancyRpartPlot(model.rpart$finalModel)
```

##Testing the Models on the Validation Set
```{r}
Validate.RF <- predict(model.RF, Validation)
confusionMatrix(Validation$classe, Validate.RF)
Validate.Rpart <- predict(model.rpart, Validation)
confusionMatrix(Validation$classe, Validate.Rpart)
```

##Model Selection
Based on the outcome of the validation tests, the Random Forests prediction model as it was 99.8% accurate in the validation test compared to the Decision Tree model at 49.3%
